Doing Better Than UCT: \\ Rational Monte Carlo Sampling in Trees

### 2011-08-02 paper/rmcts/paper.tex
paper: limit derivation draft
derivations

\begin{figure}[t]
  \centering
  \includegraphics[scale=1.2]{onelevel-tree.pdf}
  \caption{Simple regret in MAB: the best arm}
  \label{fig:onelevel-tree}
\end{figure}

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.64]{onelevel-64.pdf}
  \caption{Simple regret in MAB: UCB vs. $\varepsilon$-greedy vs. Random}
  \label{fig:onelevel-64}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includegraphics[scale=1.2]{twolevel-tree.pdf}
  \caption{MCTS: a path to the best arm}
  \label{fig:twolevel-tree}
\end{figure}

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.64,trim=0pt 0pt 0pt 24pt,clip]{twolevel-64.pdf}
  \caption{MCTS: UCT vs. GRT vs. GCT vs. Random}
  \label{fig:twolevel-64}
\end{figure}

@@

@@

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.64,trim=0pt 0pt 0pt 0pt,clip]{twolevel-8-16-32-64.pdf}
  \caption{Monte carlo tree search, varying number of arms}
  \label{fig:twolevel-8-16-32-64}
\end{figure}

@@

%\begin{figure}[t]
%  \begin{minipage}[b]{0.5\linewidth} \centering
%    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
%    0pt,clip]{sailing-size=8-factor=_8438.pdf}
%    {a. varying number of samples} 
% \end{minipage}
%  \begin{minipage}[b]{0.5\linewidth} \centering
%    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
%    0pt,clip]{sailing-size=8-nsamples=185.pdf}
%    {b. varying exploration factor}
%  \end{minipage}
%  \caption{The sailing domain, a smaller lake}
%  \label{fig:sailing-8}
%\end{figure}
%
%\begin{figure}[t]
%  \begin{minipage}[b]{0.5\linewidth} \centering
%    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
%    0pt,clip]{sailing-size=13-factor=_5625.pdf}
%    {a. varying number of samples}
%  \end{minipage}
%  \begin{minipage}[b]{0.5\linewidth} \centering
%    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
%    0pt,clip]{sailing-size=13-nsamples=240.pdf}
%    {b. varying exploration factor}
%  \end{minipage}
%  \caption{The sailing domain, a larger lake}
%  \label{fig:sailing-13}
%\end{figure}

@@

A bound on the simple regret $R_\varepsilon$ of an
$\varepsilon$-greedy policy follows from substituting
(\ref{eq:pr-epsilon-greedy}) into (\ref{eq:simple-regret}):

\begin{eqnarray}
\IE[R_\varepsilon]&\le&\sum_{j=1}^K\Delta_j\left(\mu + \frac {4\sqrt e}
{\Delta_j^2}\right)e^{-\Delta_j^2\mu/8}\nonumber\\
&\le&K\left(\mu + \frac {4\sqrt e}{\Delta_{\min}}\right)e^{-\Delta_{\min}^2\mu/8}
\end{eqnarray}

\subsubsection{UCB}

Regret $R_{UCB}$:

\subsubsection{UVB}

Regret $R_{UVB}$:

\subsection{Asymptotic bounds for $\varepsilon$-greedy policies}

>>>>

@@

@@

\subsection{Simple regret in multi-armed bandits}

\begin{figure}[t]
  \begin{minipage}[c]{0.5\linewidth}
    \centering
    \includegraphics[scale=1.0]{onelevel-tree.pdf}\\
    \vspace{4em}
    a. search tree
  \end{minipage}
  \begin{minipage}[c]{0.5\linewidth}
    \centering
    \includegraphics[scale=0.5]{flat-trilevel-k=64-uqb=8.pdf}\\
    b. regret vs. number of samples
  \end{minipage}
  \label{fig:mab-simple-regret}
  \caption{Simple regret in MAB}
\end{figure}

@@

\begin{figure}
  \begin{minipage}[c]{0.5\linewidth}
    \centering
    \includegraphics[scale=0.8]{twolevel-tree.pdf}\\
    a. search tree
  \end{minipage}
  \begin{minipage}[c]{0.5\linewidth}
    \centering
    \includegraphics[scale=0.4]{tree-identity-k=16-uqb=8.pdf}\\ 
    b. 16 arms \\
    \vspace{1em}
    \includegraphics[scale=0.4]{tree-identity-k=64-uqb=8.pdf} \\
    c. 64 arms
 \end{minipage}
  \label{fig:mcts-regret}
  \caption{MCTS: a path to the best arm}
\end{figure}

@@

\begin{figure}[t]
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[scale=0.45]{costs-size=6-group=median.pdf}\\
    a. median cost
  \end{minipage}
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[scale=0.45]{costs-size=6-group=minimum.pdf}\\
    b. minimum cost
  \end{minipage}
  \caption{The sailing domain, $6\times 6$ lake, cost vs. number of playouts}
  \label{fig:sailing-cost-vs-nsamples}
\end{figure}

\begin{figure}[t]
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[scale=0.45]{costs-size=6-nsamples=199.pdf}\\
    a. 199 playouts\\
    \vspace{1em}
    \includegraphics[scale=0.45]{costs-size=6-nsamples=397.pdf}\\
    b. 397 playouts\\
  \end{minipage}
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[scale=0.45]{costs-size=6-nsamples=793.pdf}\\
    c. 793 playouts\\
    \vspace{1em}
    \includegraphics[scale=0.45]{costs-size=6-nsamples=1585.pdf}\\
    d. 1585 playouts\\
  \end{minipage}
  \caption{The sailing domain, $6\times 6$ lake, cost vs. factor}
  \label{fig:sailing-cost-vs-factor}
\end{figure}

\begin{figure}[t]
  \begin{minipage}[b]{0.333\linewidth}
    \centering
    \includegraphics[scale=0.35]{costs-size=3-nsamples=397.pdf}\\
    a. $3\times 3$ lake
  \end{minipage}
  \begin{minipage}[b]{0.333\linewidth}
    \centering
    \includegraphics[scale=0.35]{costs-size=6-nsamples=397.pdf}\\
    b. $6\times 6$ lake
  \end{minipage}
  \begin{minipage}[b]{0.333\linewidth}
    \centering
    \includegraphics[scale=0.35]{costs-size=10-nsamples=397.pdf}\\
    b. $10\times 10$ lake
  \end{minipage}
  \caption{The sailing domain, 397 samples, cost vs. factor}
  \label{fig:sailing-lake-size}
\end{figure}

\begin{figure}[t]
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[scale=0.45]{rcq-size=6-nsamples=199.pdf}\\
    a. 199 playouts\\
    \vspace{1em}
    \includegraphics[scale=0.45]{rcq-size=6-nsamples=397.pdf}\\
    b. 397 playouts\\
  \end{minipage}
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[scale=0.45]{rcq-size=6-nsamples=793.pdf}\\
    c. 793 playouts\\
    \vspace{1em}
    \includegraphics[scale=0.45]{rcq-size=6-nsamples=1585.pdf}\\
    d. 1585 playouts\\
  \end{minipage}
  \caption{The sailing domain, log vs. sqrt, $6\times 6$ lake}
  \label{fig:sailing-cost-vs-factor}
\end{figure}
%
%\begin{figure}[t]
%  \begin{minipage}[b]{0.5\linewidth} \centering
%    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
%    0pt,clip]{sailing-size=13-factor=_5625.pdf}
%    {a. varying number of samples}
%  \end{minipage}
%  \begin{minipage}[b]{0.5\linewidth} \centering
%    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
%    0pt,clip]{sailing-size=13-nsamples=240.pdf}
%    {b. varying exploration factor}
%  \end{minipage}
%  \caption{The sailing domain, a larger lake}
%  \label{fig:sailing-13}
%\end{figure}

@@

A bound on the simple regret $r_\varepsilon$ of an
$\varepsilon$-greedy policy follows from substituting
(\ref{eq:pr-epsilon-greedy}) into (\ref{eq:simple-regret}):

\begin{eqnarray}
  \IE r_\varepsilon&\le&\sum_{j=1}^K\Delta_j\left(\mu + \frac {4\sqrt e}
{\Delta_j^2}\right)e^{-\Delta_j^2\mu/8}\nonumber\\
&\le&K\left(\mu + \frac {4\sqrt
    e}{\Delta_{\min}}\right)e^{-\Delta_{\min}^2\mu/8}
\label{eq:deriv-epsilon-greedy-b}
\end{eqnarray}

\subsubsection{UCB($\alpha$)}

$n_b$ pulls of the best arm and $n_j$ pulls of the $j$th arm are
achieved simultaneously when either

\begin{equation}
\sqrt {\frac {\alpha \log n} {n_j-1}} \ge 1 + \sqrt { \frac {\alpha \log n} {n_b}}
\end{equation}

or

\begin{equation}
\sqrt {\frac {\alpha \log n} {n_j}} \le 1 + \sqrt { \frac {\alpha \log n} {n_b - 1}}
\label{eq:deriv-ucb-alpha-a}
\end{equation}

Inequality (\ref{eqn:deriv-ucb-alpha-a}) can be used to derive a lower
bound on the number of pulls of a non-optimal arm,  and, consequently, the upper
bound on the probability of selecting a non-optimal arm:

\begin{eqnarray}
\frac {\alpha \log n} {n_j}&\le& 1 + 2  \sqrt {\frac {\alpha \log n}
  {n_b-1}} + \frac {\alpha \log n} {n_b-1}\nonumber\\
n_j &\ge& \frac {\alpha \log n} {1+2 \sqrt {\frac {\alpha \log n}
    {n_b-1}} + \frac {\alpha \log n} {n_b-1}}\nonumber\\
    &\ge& \frac {\alpha \log n} {4} \mbox{ for } \alpha \log n \le n_b-1
\end{eqnarray}

The probability that the $j$th arm will be chosen can be bounded based
on the Chernoff inequality and the union bound:
\begin{equation}
P_j \le 2 \exp\left(-\frac {\alpha \Delta_j^2 \log n} 8 \right) = 2n^{\frac {-\alpha \Delta_j^2} 8}
\end{equation}

The expected simple regret of an UCB($\alpha$) algorithm is thus
bounded by the following inequality:
\begin{equation}
\IE r_{ucb} \le 2\sum_{j=1}^K \Delta_jn^{\frac {-\alpha \Delta_j^2} 8}
\label{eq:deriv-ucb-alpha-b}
\end{equation}

\cite{Bubeck.pure} provides for UCB($\alpha$) a similar bound,
polynomial in the number of pulls $n$.

\subsection{Asymptotic regret analysis}

Bounds (\ref{eq:deriv-epsilon-greedy-b}, \ref{eq:deriv-ucb-alpha-b}) where
derived ignoring dependency of the allocation on expected values
of non-optimal arms. Derivation of better bounds seems complicated; instead, insights
can be obtained by analysing  asymptotic behavior of each of the
algorithms in assumption that the sample mean approaches the expected
value.

\subsubsection{UCB($\alpha$)}
